\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}

### 2023-10-24 manuscript/manuscriptRW.Rmd
last changes before sent for internal round
add revisions from Harald

- Mats Mahnken$^8$,
- Paola	Mairota$^9$,
- Ján Merganič$^{10}$,
- Katarina Merganičová$^{11,12}$,
- Tobias Mette$^{13}$,
- Marco Mina$^{14}$,
- Xavier Morin$^{15}$

@@

Most of the models overestimate recruitment levels, which is only
compensated in some models by high simulated mortality rates in the
early stages of individual trees dynamics. Simulated species diversity
of recruitment matches the observed ranges. Models simulating higher
species diversity at the stand level do not present higher species
diversity in the recruitment layer. The effect of light availability on
recruitment levels is captured better than the effect of temperature and
soil moisture, but patterns are not consistent across models. Increasing
complexity in the tree regeneration modules of the models is not related
to higher accuracy of simulated tree recruitment. Furthermore,
individual model design is more important than scale (stand, landscape,
global) and approach (empirical, process-based) for accurately capturing
tree regeneration.

Despite the mismatches between simulation results and data, it is
remarkable that most models capture the essential features of the highly
complex process of tree regeneration, while not having been
parameterized with such data. We conclude that much can be gained by
evaluating and refining the modeling of regeneration processes. This has
the potential to render long-term projections of forest dynamics under
changing environmental conditions that are much more robust.

@@

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics have been developed over the past decades considering
the impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
less studied process that is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. This presents an important
research gap, particularly in the context of climate-induced forest
disturbances and forest resilience.

@@

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree regeneration e.g. in
the context of global change scenarios, and (3) identify the most
important processes that are shaping ecological patterns. Given the
current strategies that are used in models of forest dynamics to
represent tree regeneration, model behavior often is prone to problems,
such as very high levels of tree regeneration that necessitate excess
mortality at early stages of tree life to simulate correct stand
structure and composition [@kroiss2015]. Also, inadequately high species
diversity in tree regeneration may be simulated, which is characteristic
of "classical" forest gap models [@gutiérrez2016], at least as long as
the simulation set-up comprises a multitude of species. Some models use
calibration against local canopy-level data to constrain simulated
regeneration levels, which is likely to hamper the general applicability
of these models e.g. under scenarios of climate change. Furthermore,
correctly capturing the species composition in tree regeneration as a
function of the presence of seed trees in the canopy is often a
particular challenge, potentially leading to unrealistic successional
drift in the model, which must be corrected by factors that are hard to
parameterize [e.g., @lischke2006]. A related issue is the excessive
reduction of species diversity due to positive feedback effects, such
that eventually just single-species stands remain [@meier2011]. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm in height
or even seeds) and track tree development in a process-oriented manner
by considering a multitude of ecological influences, rather than
emphasizing tree regeneration into a larger size class [e.g.,
@wehrli2006].

Modeling tree regeneration processes is challenging, and even in
empirical ecology, it has not received much attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that cover all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Data on forest regeneration are often
fragmented, which constitutes a major problem for model building
[@clark1999a]. For example, monitoring on permanent plots (such as
National Forest Inventories) often measures tree regeneration
[@zell2019], however with a design that captures the rate after the
stand initiation phase [@hallsby2015]. Data from permanent plots,
although available across regions, are also highly heterogeneous,
rendering their use difficult in a modeling context [@käber2021], and
targeted studies to measure tree regeneration on experimental sites are
often limited in spatial extent, species studied, or the sub-set of
processes that are targeted [@berdanier2016; @collet2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. In the present
study, for the first time, a large number of forest models commonly used
to assess forest dynamics under climate change are evaluated against a
continental-scale, multi-species harmonized dataset on tree recruitment
[@käber2023]. By tree recruitment, we refer to the passing of trees
across a specific diameter threshold ("ingrowth"). We included models
that are based on a range of "philosophies", from models operating at
the stand to the global scale as well as the range of models from
empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving and constraining this process --- including some that are not
incorporated explicitly in most models, such as deer browsing --- we do
not aim for a detailed statistical evaluation of each model. Instead, we
aim to evaluate the general regeneration patterns and magnitudes
simulated by the models, and benchmark the simulated regeneration niche
of multiple species against empirical data along a wide environmental
gradient of temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree recruitment levels,
initial tree species diversity and mortality in the recruitment? (2) Do
model traits explain differences in model performance? (3) How well do
the models capture total recruitment and individual species regeneration
niches across environmental gradients of light availability,
temperature, and soil moisture? We evaluate the performance of
regeneration models by analyzing simulated data alongside observed data.
Specifically, we examine tree species diversity, mortality rates in tree
recruitment, and recruitment levels. Then, we link model traits (e.g.,
complexity, scale) to performance, and we assess whether modeled total
recruitment or individual species regeneration niches (i.e.,
proportional recruitment along an environmental gradient) align with
observational data.

@@

Fifteen models of forest dynamics (including two models featuring model
variants) were used to simulate forest dynamics at the stand, landscape
or global scale (Table \@ref(tab:tableModels)). The approaches used for
model construction and their origin differ strongly, with most of the
models featuring a largely "process-based" approach, whereas two models
are based on formulations derived from the statistical analysis of
inventory data (SIBYLA and xComp). Some of the models largely rely on
the original approach underlying forest gap models (e.g., ForClim 1),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree regeneration processes.

The regeneration models can be differentiated into 'regeneration' and
'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
models* include processes such as flowering and pollination, seed
production, seed dispersal, germination, and seedling growth, which
ultimately lead to the simulated number of established trees.
*Recruitment models*, in contrast, introduce a number of new trees with
certain characteristics such as biomass or diameter, without explicitly
considering earlier development processes. We can further distinguish
models that feature a feedback in the regeneration, i.e., where the
level and species composition of regeneration is influenced by the
existing adult tree community via the production of seeds, seedlings or
saplings, from models that do not contain such a feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes that are not treated explicitly. This latter approach reduces
model complexity but comes at the cost of blurring process
representation.

In the models used here, the overall complexity in the regeneration
modules varies considerably. Following @bugmann2022, we can classify
seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
TreeMig) as having rather high complexity in their regeneration modules
(mean regeneration formulation complexity across all processes \>0,
Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
used here with two alternative variants of regeneration. ForClim variant
1 [@bugmann1996] is based on a regeneration module that adheres closely
to the concept introduced by @botkin1972 which is an environmental
filter considering species specific thresholds of light availability and
climatic variables. Whereas ForClim variant 11 is adopting a slightly
more complex approach where individual species properties and their
relation to the environment are represented gradually along with the
relative suitability of different species (i.e., competition among tree
species) [@huber2020]. These two model variants allow us to evaluate a
more process-based and complex module (variant 11) against a simple
module (variant 1), while the rest of the model structure is identical.
Similarly, the two ForCEEPS variants allow us to isolate the importance
of the canopy feedback (i.e., simulated actual composition and relative
abundance of species in the plot) via seed trees for the quantity and
quality (e.g., diversity and composition) of simulated regeneration, as
one variant includes this feedback whereas the other does not.

@@

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the framework of the EuFoRIa network
[@euforia2019; @käber2023] (Figure 1). The environmental coverage and
the variety of forests represented by this data is unprecedented within
Europe. These forests have been unmanaged for at least 10 years prior to
their designation as reserves, and most of them provide time series of
natural forest dynamics over multiple decades and up to ca. 80 years.
The census periods range from 3 to 37 years, with an average of 14
years. The data provide information at the tree level, thus allowing for
the sequential comparison of processes such as individual-tree
regeneration and death between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality, slope and aspect. Each of these sites featured at least
two consecutive measurements. Recruitment minimum size thresholds for
these sites differed between stem diameters of 0 and 10 cm. For our
study, we defined two datasets where one included 165 sites with a
diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds from 10 cm. Therefore we had in total 200
sites with observations of newly established trees above 10 cm and 165
sites above 7cm. In the results section, any result from observed data
at 7 cm threshold originated from the 165 sites.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2023 to provide recruitment
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Recruitment rates per site,
sample and decade ranged from zero to 1246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this dataset and the detailed regeneration information,
cf. @käber2023.

@@

The input variables were collected from different data sources and
aggregated to be adapted to the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data were slightly different in each model, based
on the approach that best suited the model (Table
\@ref(tab:tableModels)). Soil quality data were provided as continuous
values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
The protocol also provided the elevation, slope and aspect for each of
the 200 sites [@asterscienceteam2019], but no other spatial information
such as coordinates, with the exception of iLand and aDGVM2, which
required blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a background disturbance to increase
chances of shade-intolerant species to establish. The simulations were
set up to sample species-specific recruitment rates per decade and per
ha in the equilibrium state of the model, typically entailing a
"spin-up" run (as we did not provide any forest data). The modeling
teams decided on the simulated area and how they derived these samples.
The exact length of the simulation was also decided by the modeling
teams (Table \@ref(tab:tableModels)). Further details on how each
modeling team prepared the simulations and the outputs are available in
Supplementary Material 1.

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree recruitment under comparable
and near-equilibrium conditions. This entails the assumption that (1)
the observations from the forest reserves reflect no traces of forest
management, and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of the regeneration process (i.e., the number of recruited
trees per area and per unit of time) were of interest.

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig), the species
simulated beyond the eleven species were lumped in a category "others".
In three models, fewer than these eleven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
sylvestris* are parameterised. xComp simulations did not consider
*Fraxinus excelsior* due to a deprecated species parameterisation. In
LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
parameterized and therefore these species could not be included in the
simulation. Finally, the model aDGVM2 does not simulate individual
species. Rather, community assembly processes and trait filtering
generate plant communities that are adjusted to the biotic and abiotic
conditions, and the simulated plants can be classified into ecological
strategies based on their trait values in a post-processing step.

@@

Mortality in tree recruitment was assessed based on the ratio of
recruitment between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
to estimate whether the ratio of recruitment between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate.
The Reineke self-thinning rule is usually calculated for even-aged,
single species stands and assumes a fixed relationship between the
number of stems and the quadratic mean diameter in fully stocked pure
stands. The value used in our comparisons was 1.77 (i.e., we expect stem
numbers at 7 cm to be 77% higher than at 10 cm), calculated using
Equation \ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using paired t-tests.

Total recruitment and the regeneration niches of the individual species
were evaluated across the environmental gradients of light availability,
temperature, and soil moisture. Stand basal area (the basal area of all
the trees in each sample including tree regeneration) was used as a
proxy for light availability at the forest floor, the annual degree-day
sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
and the climatic water balance as a proxy for soil moisture
[@speich2019]. Recruitment values of each species were calculated as the
mean across the 200 samples per site. The observed data were modeled
using a Generalized Additive Model [@wood2011] with a negative binomial
distribution and restricted maximum likelihood to better understand the
relationship between the environmental gradients and the levels of
observed regeneration, relative to the simulation results per model.

@@

## Regeneration levels

@@

Both the observations and the simulated data had no recruitment in some
samples and at some sites (for details, cf. Table S1). The observed data
had 4% of the samples with no recruitment. Only three models had a
larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
models always simulated recruitment for both the 7 and 10 cm threshold
(xComp and TreeMig), i.e., they did not feature any zero values. The
other ten models had a very low percentage of samples with no
recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
of zeros compared to the observations.

## Tree species diversity of regeneration

@@

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure 4A, left and center; cf. Figure
S3), but a characteristic feature was that the simulations had a much
lower variability of diversity compared to observations. Models
overpredicting species diversity included ForClim1, ForClim 11, FORMIND
and TreeMig (Figure 4A, left). Several models underpredicted species
diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A, right).

>>>>

- Paola	Mairota$^8$,
- Ján Merganič$^{9}$,
- Katarina Merganičová$^{10,11}$,
- Tobias Mette$^{12}$,
- Marco Mina$^{13}$,
- Xavier Morin$^{14}$
- Mats Nieberg,$^{15}$,

@@

Most of the models overestimate recruitment levels, which is compensated
only in some models by high simulated mortality rates in the early
stages of individual tree dynamics. Simulated species diversity of
recruitment generally matches observed ranges. Models simulating higher
stand-level species diversity do not feature higher species diversity in
the recruitment layer. The effect of light availability on recruitment
levels is captured better than the effects of temperature and soil
moisture, but patterns are not consistent across models. Increasing
complexity in the tree regeneration modules is not related to higher
accuracy of simulated tree recruitment. Furthermore, individual model
design is more important than scale (stand, landscape, global) and
approach (empirical, process-based) for accurately capturing tree
regeneration.

Despite the mismatches between simulation results and data, it is
remarkable that most models capture the essential features of the highly
complex process of tree regeneration, while not having been
parameterized with such data. We conclude that much can be gained by
evaluating and refining the modeling of tree regeneration processes.
This has the potential to render long-term projections of forest
dynamics under changing environmental conditions much more robust.

@@

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics have been developed over the past decades considering
the impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is much
less studied and is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. This presents an important
research gap, particularly in the context of climate-induced forest
disturbances and forest resilience.

@@

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree regeneration e.g. in
the context of global change scenarios, and (3) identify the most
important processes that are shaping vegetation patterns. Given the
strategies that are used in models of forest dynamics to represent tree
regeneration, their behavior often is prone to problems, such as very
high levels of tree regeneration that necessitate excess mortality at
early stages of tree life to simulate correct stand structure and
composition [@kroiss2015]. Also, inadequately high species diversity in
tree regeneration may be simulated, which is characteristic of
"classical" forest gap models [@gutiérrez2016], at least as long as the
simulation set-up comprises a multitude of species. Some models use
calibration against local canopy-level data to constrain simulated tree
regeneration, which is likely to hamper the general applicability of
these models e.g. under scenarios of climate change. Furthermore,
correctly capturing the species composition of tree regeneration as a
function of the presence of seed trees in the canopy is often a
particular challenge, potentially leading to unrealistic successional
drift in the model, which must be corrected by factors that are hard to
parameterize [e.g., @lischke2006]. A related issue is the excessive
reduction of species diversity due to positive feedback effects, such
that eventually just single-species stands remain [@meier2011]. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm in height,
or even from seed) and track tree development in a process-oriented
manner by considering a multitude of ecological influences, rather than
emphasizing tree recruitment into a larger size class [e.g.,
@wehrli2006].

Modeling tree regeneration processes is challenging, and even in
empirical ecology, it has not received much attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that cover all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Data on forest regeneration are often
fragmented, which constitutes a major problem for model building
[@clark1999a]. For example, monitoring on permanent plots (such as
National Forest Inventories) often measures the ingrowth of new trees
into a specific size class [@zell2019], however with a design that
captures the rate after the stand initiation phase [@hallsby2015]. Data
from permanent plots, although available across regions, are also highly
heterogeneous, rendering their use difficult in a modeling context
[@käber2021]. Lastly, targeted studies to measure tree regeneration on
experimental sites are often limited in spatial extent, species studied,
or the sub-set of processes that are investigated [@berdanier2016;
@collet2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. In the present
study, for the first time, a large number of forest models commonly used
to assess forest dynamics under climate change are evaluated against a
continental-scale, multi-species harmonized dataset on tree recruitment
[@käber2023]. By tree recruitment, we refer to the passing of trees
across a specific diameter threshold ("ingrowth"). We included models
that are based on a range of "philosophies", from models operating at
the stand to the global scale as well as the range of models from
empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving and constraining this process --- including some that are not
incorporated explicitly in most models, such as deer browsing --- we do
not aim for a detailed statistical evaluation of each model. Instead, we
aim to evaluate the general recruitment patterns and magnitudes
simulated by the models, and benchmark the simulated regeneration niche
of multiple species against empirical data along a wide environmental
gradient of temperature, moisture, and light availability [@grubb1977].

More specifically, we aim to answer the following questions: (1) Are
models of forest dynamics capturing accurately tree recruitment levels,
initial tree species diversity and mortality in the recruitment? (2) Do
model traits explain differences in model performance? (3) How well do
the models capture total recruitment and the regeneration niches of
individual species across environmental gradients of light availability,
temperature, and soil moisture?

We evaluate the performance of the models by analyzing simulated data
alongside observed data. Specifically, we examine tree species
diversity, mortality rates in tree recruitment, and recruitment levels.
Then, we link model traits (e.g., complexity, scale) to performance, and
we assess whether modeled total recruitment or individual species
regeneration niches (i.e., proportional recruitment along an
environmental gradient) align with observational data.

@@

Fifteen models of forest dynamics (including two models featuring model
variants) developed for the stand, landscape or global scale were used
to simulate forest dynamics (Table \@ref(tab:tableModels)). The
approaches used for model construction and their origin differ strongly,
with most of the models featuring a largely "process-based" approach,
whereas two models are based on formulations derived from the
statistical analysis of inventory data (SIBYLA and xComp). Some of the
models largely rely on the original approach underlying forest gap
models (e.g., ForClim 1), whereas some are based on plant
ecophysiological processes (e.g., FORMIND and iLand). The different
approaches underlying the models have strongly influenced the
formulation of tree regeneration processes.

The models can be differentiated into 'regeneration' and 'recruitment'
approaches [@vanclay1994; @konig_tree_2022]. *Regeneration models*
include processes such as flowering and pollination, seed production,
seed dispersal, germination, and seedling growth, which ultimately lead
to the simulated number of established trees. *Recruitment models*, in
contrast, introduce a number of new trees with certain characteristics
such as biomass or diameter, without explicitly considering earlier
development processes. We can further distinguish models that feature a
feedback from stand properties to simulated recruitment, i.e., where the
level and species composition of recruitment is influenced by the adult
tree community via the production of seeds (or seedlings/saplings), from
models that do not contain such feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes that are not treated explicitly. This latter approach reduces
model complexity but comes at the cost of blurred process
representation.

In the models used here, the overall complexity in the regeneration
modules varies considerably. Following @bugmann2022, we can classify
seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
TreeMig) as having rather high complexity in their regeneration modules
(mean regeneration formulation complexity across all processes \>0,
Table \@ref(tab:tableModels)), whereas the other eight models feture
relatively low complexity. Two models, ForClim and ForCEEPS, were used
here with two alternative variants of regeneration. ForClim variant 1
[@bugmann1996] is based on a regeneration module that adheres closely to
the concept introduced by @botkin1972 which is based on environmental
filters considering species-specific thresholds of light availability
and climatic variables. In contrast, ForClim variant 11 is adopting a
slightly more complex approach where individual species properties and
their relation to the environment are represented gradually along with
the relative suitability of different species (i.e., competition among
tree species) [@huber2020]. These two model variants allow us to
evaluate a more process-based and complex module (variant 11) against a
simple module (variant 1), while the rest of the model structure is
identical. Similarly, the two ForCEEPS variants allow us to isolate the
importance of the canopy feedback (i.e., simulated actual composition
and relative abundance of species in the plot) via seed trees for the
quantity and quality (e.g., diversity and composition) of simulated
regeneration, as one variant includes this feedback whereas the other
does not.

@@

Recruitment data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel network of sites in forest reserves that represent the range of
environmental gradients in temperature and precipitation in Central
Europe as compiled in the EuFoRIa network [@euforia2019; @käber2023]
(Figure 1). The environmental coverage and the variety of forests
represented by this data is unprecedented within Europe. These forests
have been unmanaged for at least 10 years prior to their designation as
reserves, and most of them provide time series of natural forest
dynamics over multiple decades and up to ca. 85 years. The census
periods range from 3 to 37 years, with an average of 14 years. The data
provide information at the tree level, thus allowing for the sequential
comparison of processes such as individual-tree recruitment and death
between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation, so as to be representative of the environmental
variation contained in the data. This was achieved by applying k-means
clustering to define 200 clusters of plots from the original set of 869
plots along the environmental dimensions of temperature, climatic water
balance, soil quality, slope and aspect. Each of these sites featured at
least two consecutive measurements. Recruitment size thresholds ranged
from stem diameters at breast height (dbh) from 0 to 10 cm. For our
study, we defined two datasets, where one included 165 sites with a
diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm. Therefore we had in
total 200 sites with observations of newly recruited trees above 10 cm
and a subset of 165 sites with recruitment data above 7 cm. In the
results section, any observations from the data at the 7 cm threshold
originated from 165 sites.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2023 to provide recruitment
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Recruitment rates per site,
sample and decade ranged from zero to 1,246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this dataset and the detailed recruitment information,
cf. @käber2023.

@@

The input variables were collected from different data sources and
aggregated to be suitable for the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data were slightly different in each model, based
on the approach that best suited the model (Table
\@ref(tab:tableModels)). Soil quality data were provided as continuous
values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
The protocol also provided the elevation, slope and aspect for each of
the 200 sites [@asterscienceteam2019], but no other spatial information
such as coordinates, with the exception of iLand and aDGVM2, which
required blurred coordinates to derive highly detailed soil data.

The simulations were run in the absence of natural disturbances. Only
the model LPJ-GUESS had to include a "background" disturbance
probability to increase chances of shade-intolerant species to
establish. The simulations were set up to sample species-specific
recruitment rates per decade and per ha in the equilibrium state of the
model, typically entailing a "spin-up" run (as we did not provide any
forest data). The modeling teams decided on the simulated area and how
they derived these samples. The exact length of the simulation was also
decided by the modeling teams (Table \@ref(tab:tableModels)). Further
details on how each modeling team prepared the simulations and the
outputs are available in Supplementary Material 1.

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree recruitment under comparable
and near-equilibrium conditions. This entails the assumption that (1)
the observations from the forest reserves reflect no traces of forest
management, and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of recruitment (i.e., the number of recruited trees per area
and per unit of time) were of interest.

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L.,* and *Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig), and the species
simulated beyond the eleven species were lumped in a category "others".
In three models, fewer than these eleven species were simulated (4C,
xComp, LPJ-GUESS) (Table \@ref(tab:tableModels)). In 4C only *Fagus
sylvatica*, *Picea abies*, *Betula* spp., *Quercus* spp. and *Pinus
sylvestris* are parameterised. xComp simulations did not consider
*Fraxinus excelsior* due to a deprecated species parameterisation. In
LPJ-GUESS, *Acer pseudoplatanus* and *Alnus glutinosa* are not
parameterized and therefore these species could not be included in the
simulation. Finally, the model aDGVM2 does not simulate individual
species. Rather, community assembly processes and trait filtering
generate plant communities that are adjusted to the local biotic and
abiotic conditions, and the simulated plants can be classified into
ecological strategies based on their trait values in a post-processing
step.

@@

Mortality in tree recruitment was assessed based on the ratio of
recruitment between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
to estimate whether the ratio of recruitment between these two
thresholds was above or below the expected theoretical rate. The Reineke
self-thinning rule is usually calculated for even-aged, single species
stands and is based on a fixed relationship between the number of stems
and the quadratic mean diameter in fully stocked pure stands. The value
used in our comparisons was 1.77 (i.e., we expect stem numbers at 7 cm
to be 77% higher than at 10 cm), calculated using Equation
\ref{eqReineke}.

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using paired t-tests.

Total recruitment and the regeneration niches of the individual species
were evaluated across the gradients of light availability, temperature,
and soil moisture as captured in the data. Stand basal area (the basal
area of all the trees in each sample above the recruitment threshold)
was used as a proxy for light availability at the forest floor, the
annual degree-day sum [@allen1976; @fischlin1995] as a proxy for growing
season warmth, and the climatic water balance as a proxy for soil
moisture [@speich2019]. Recruitment values of each species were
calculated as the mean across the 200 samples per site. The observed
data were modeled using a Generalized Additive Model [@wood2011] with a
negative binomial distribution and restricted maximum likelihood to
better understand the relationship between the environmental gradients
and the levels of observed regeneration, relative to the simulation
results per model.

@@

## Recruitment levels

@@

Both the observations and the simulated data had no recruitment in some
samples and at some sites (for details, cf. Table S1). The observed data
had 4% of the samples with no recruitment. Only three models had a
larger proportion of no recruitment (4C, Landis-II and aDVM2). Two
models always simulated recruitment for both the 7 and 10 cm threshold
(xComp and TreeMig), i.e., they did not feature any zero values. The
other ten models had a very low percentage of samples with no
recruitment (0.01% - 2.39%), i.e., they had distinctly fewer occurrences
of zeros than the observations.

## Tree species diversity of recruitment

@@

For both the observed and the simulated data, and both recruitment
thresholds, species diversity in the recruitment compared to stand-level
diversity did not feature clear patterns across the models (Figures S1
and S2). Most models captured reasonably well or overpredicted species
diversity at the stand level (Figure 4A, left and center; cf. Figure
S3), but a characteristic feature was that the simulations had a much
lower variability of diversity compared to observations. Models
overpredicting stand-level species diversity included ForClim1, ForClim
11, FORMIND and TreeMig (Figure 4A, left). Several models underpredicted
species diversity, i.e., LPJ-GUESS, ForCEEPS(f) and 4C (Figure 4A,
right).
